---
title: "Public Awareness Of And Trust In Official Statistics, Northern Ireland 2022"
lang: "en"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    code_download: no
    css: "style.css"
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "../outputs/",
      envir = globalenv()
    )
  })
---

<div id = "cookie-banner"></div>
<script src = "cookies_script.js"></script>

```{r settings, include=FALSE}
# this chunk sets chunk options for all chunks in this file
knitr::opts_chunk$set(
  message = FALSE, echo = FALSE, warning = FALSE, out.width = "100%", fig.height = 4
)

library(here)
source(paste0(here(), "/code/data_prep.R"))
```


<div data-nics-theme="`r nics_theme`">

`r f_banner()` 

<div id="main" role="main">


[Back to full report](public-awareness-of-and-trust-in-official-statistics-northern-ireland-`r current_year`.html)

## Appendix C: Survey Questions

### Qu1

"Before being contacted about this survey had you heard of NISRA, the Northern Ireland Statistics and Research Agency?"

1. Yes
1. No

_If answered 'yes' to Qu1 then_

<div style = "margin-left: 30px">

#### Qu1c 

NISRA produces official statistics for Northern Ireland on a wide range of issues. 

I am going to list a number of these and want you to tell me if you knew before today that this information was produced by NISRA statisticians. 

```{r PCOS1c}

pcos1 <- data.frame(
  Question = outputs) %>%
  mutate(`Aware produced by NISRA?` = rep("Yes/No", nrow(.)))

kbl(pcos1) %>%
  kable_styling()

```

</div>

_If answered 'no' to Qu1 then_

<div style = "margin-left: 30px">

#### Qu1d 

I am going to list a number of these and want you to tell me if you have heard of them or not.

```{r PCOS1d}
kbl(pcos1) %>%
  kable_styling()
```

</div>

### Qu2

I will name a list of institutions. For each, please indicate whether you tend to trust it or tend not to trust it.

<div style = "margin-left: 30px">

#### Qu2a

How much do you trust the Civil Service?

1. Trust a great deal
1. Tend to trust
1. Tend to distrust
1. Distrust greatly

#### Qu2b

How much do you trust the NI Assembly?

1. Trust a great deal
1. Tend to trust
1. Tend to distrust
1. Distrust greatly

#### Qu2c

How much do you trust the media?

1. Trust a great deal
1. Tend to trust
1. Tend to distrust
1. Distrust greatly

#### Qu2d

How much do you trust NISRA?

1. Trust a great deal
1. Tend to trust
1. Tend to distrust
1. Distrust greatly

</div>

### Qu3

Personally, how much trust do you have in statistics produced by NISRA? 

1. Trust them greatly
1. Tend to trust them
1. Tend not to trust them
1. Distrust them greatly

### Qu4

To what extent do you agree or disagree with the following statement ‘Statistics produced by NISRA are important to understand Northern Ireland?’

1.Strongly agree
1. Tend to agree
1. Tend to disagree
1. Strongly disagree

### Qu5

To what extent do you agree or disagree with the following statement ‘I believe that the statistics produced by NISRA are free from political interference’

1. Strongly agree
1. Tend to agree
1. Tend to disagree
1. Strongly disagree

### Qu6

To what extent do you agree or disagree with the following statement: 'I believe that personal information that is provided to NISRA will be kept confidential.'

1. Strongly agree
1. Tend to agree
1. Tend to disagree
1. Strongly disagree

<!-- closing main div -->
</div>

`r f_footer()`

<!-- message to rotate screen if it is below 400px width -->
<div id="rotate-screen">
  <p style="color: white;">Please rotate your screen for optimal experience</p>
</div>


```{js}

// Function to remove map shapes from tab order, and add tabs to table of contents

function tabOrder() {
  
  // Make side menu tabbable

  var headings = document.getElementsByClassName('tocify-item list-group-item');
  
  for (let i = 0; i < headings.length; i ++) {
    headings[i].innerHTML = "<a href='#'>" + headings[i].innerHTML +  "</a>";
  }
  
};

// Instruct web browser to run above function when loaded
window.onload = tabOrder;
```


<!-- close data-nics-theme div-->
</div>
