---
title: "06 Map Sub Report"
output: html_document
knit: (function(input, ...) {
    rmarkdown::render(
      input,
       output_dir = "../demo/demo_outputs/",
      output_file = paste0(
        xfun::sans_ext(input), '_', format(Sys.time(), "%d-%m-%Y_%H%M") , '.html'
      ),
      envir = globalenv()
    )
  })
---
  
  

```{r, include=FALSE}
# this chunk sets chunk options for all chunks in this file
knitr::opts_chunk$set(
  warning = FALSE, message = FALSE, echo = FALSE
)

library(here)

source(paste0(here(), "/code/config.R"))
source(paste0(here(), "/code/demo/demo_data_prep.R"))
source(paste0(here(), "/code/demo/demo_config.R"))
```


This section includes a static map and an interactive map. If you create a map in another package and import as an image, ensure it meets accessibility requirements for an infographic.

### Static Map (h3)

This section includes a map generated in R. General principles to consider are:

* ensure contrast between text labels and background colour is at least 4.5:1
* ensure contrast between adjacent colours is at least 3:1, if they don't use borders that do
* verify using online tools like [WebAIM contrast checker](https://webaim.org/resources/contrastchecker/)
* ensure colour is not the only way of conveying meaning
* use clear and descriptive labels

Here the data labels are displayed on the map, alt text is included and the shapefile has been simplified to reduce the file size.

We have used three colours. More than three colours increases the difficulty of creating sufficient contrast.

Ensure the data underlying the maps is accessible to users who cannot perceive images. In this case, the data is in a table in a previous section. You could also make it available as a download.

Include descriptive text before or after the map.

#### Map 1: Proportion of people aged under 25 in NI by LGD

```{r map_static, echo = TRUE, warnings = FALSE, message = FALSE, fig.alt="Map of Northern Ireland shaded by percentage of under 25s living in each Local Government District. More detail on the trends are included in the text directly above or below the map", out.width="100%"}
map1_static_map <- tm_shape(df_map_data,
  # Default is km
  unit = "imperial"
) +
  tm_polygons( # Colours the map sections based on these values
    col = "lgd_young_perc",
    # Legend title
    title = "Percentage of under 25s",
    # Number of different colours to use
    n = 3,
    # Quantiles gives even sized groupings
    style = "quantile",
    # Colour palette. Execute the command tmaptools::palette_explorer() in R
    # console to see range of options
    palette = "Blues",
    # Add transparency to polygons improving contrast with text
    alpha = 0.5,
    legend.show = FALSE,
    border.col = "black"
  ) +

  tm_text( # add labels to polygons, change text colour to black and move
    # labels to fit within shapes
    "map_label",
    size = 0.8,
    col = "#000000",
    auto.placement = FALSE,
    xmod = c(0.8, 0, 0, 0.5, 0, 1.1, 0.5, 0.5, 0, 1, -1),
    ymod = c(-0.5, 0, 0.2, 0.3, 0, 0, 1.4, -0.5, 0.3, 0, -1)
  ) +
  tm_add_legend("fill",
    col = RColorBrewer::brewer.pal(3, "Blues"),
    border.col = "#454A58",
    labels = c("26.6  to 29.5", "29.6 to 31.5", "31.6 to 33.5"),
    title = "Percentage of under 25s"
  ) +
  tm_add_legend( # add manual legend as a key
    "symbol",
    col = "#00205b",
    title = "Key",
    labels = c(
      "AN = Antrim and Newtownabbey",
      "AND = Ards and North Dow",
      "B = Belfast"
    )
  ) +
  tm_format( # setting format to worldwide creates a wider rectangle shape that
    # fits width of page
    "World_wide",
    legend.position = c("left", "top")
  )

# Plotting map:
tmap_mode("plot") # Set to "view" for interactive map or "plot" for static

map1_static_map
```


### Interactive Maps (h3)

This section shows the same map produced as an interactive plot.
The figures are displayed on the map, council names show on hover and the council name and figures display on click.
The shapefile has been simplified to reduce file size and not slow page load speed.

Ensure the data underlying the maps is accessible to users who cannot perceive images. In this case, the data is in a table in a previous section. You could also make it available as a download.

The default rendering tabs through the LGDs but the focus area is not visible for keyboard users. This is an accessibility fail. Additional code has been added to remove the map from the tab order and to add alt text to the LGDs for screen reader users. See sections of code in the final code chunk (download buttons) which creates alt text and in the script section at the very end of the code to create an array of alt texts for the map and remove tabs from the LGDs.  

Include descriptive text before or after the map.

#### Map 2: Interactive shapefile map with hover text

```{r map_interactive_1, echo = TRUE, warnings = FALSE, message = FALSE, fig.alt="alt text for interactive map", fig.cap="caption for interactive map", out.width="100%"}
# Reduce the size of the shapefile using the simplify command in rmaptools.


#### interactive map ####

map2_interactive_map <- tm_shape(
  df_map_data %>%
    select(lgd2014name, lgd_young_perc, map_label_value_only),
  name = "Show data",
  # Default is km
  unit = "imperial"
) +

  tm_basemap( # Adds background map.
    "Esri.WorldGrayCanvas"
  ) +
  tm_polygons( # Colours the map sections based on these values
    col = "lgd_young_perc",
    # Number of different colours to use
    n = 3,
    # Quantiles gives even sized groupings
    style = "quantile",
    # Colour palette
    palette = "Blues",
    # Add transparency to polygons improving contrast with text
    alpha = 0.5,
    border.alpha = 1,
    legend.show = FALSE,
    popup.vars = c("Percentage of under 25s: " = "lgd_young_perc")
  ) +
  tm_add_legend("fill",
    col = RColorBrewer::brewer.pal(3, "Blues"),
    border.col = "#454A58",
    labels = c("26.6  to 29.5", "29.6 to 31.5", "31.6 to 33.5"),
    title = "Percentage of under 25s"
  ) +
  tm_view( # Re-positioning legend
    view.legend.position = c("left", "top"),
    # Re-positioning layer controls
    control.position = c("left", "bottom"),
    leaflet.options = c(
      zoomControl = FALSE,
      # Removes zoom and drag options
      dragging = FALSE,
      keyboard = FALSE,
      boxZoom = FALSE,
      doubleClickZoom = FALSE,
      scrollWheelZoom = FALSE,
      tap = FALSE,
      touchZoom = FALSE,
      zoom = 8,
      minZoom = 8,
      maxZoom = 8,
      attributionControl = FALSE
    )
  ) +
  tm_text( # Label polygons, increase font size, set colour to black (better
    # contrast) and adjust positioning so they are contained as much as
    # possible within the polygons
    "map_label_value_only",
    size = 1.2,
    col = "#000000",
    auto.placement = FALSE,
    xmod = c(3, 1, 0, 1.5, 0, 1.5, 0, 0, 0, 1, 1),
    ymod = c(0, 3, 0, 0, 0, 0, 1, 0, 0, 0, -3)
  ) +
  tm_scale_bar(width = 250) +
  tmap_mode( # Set to "view" for interactive map or "plot" for static
    "view"
  )

map2_interactive_map

# please also refer to the "alt text for maps" chunk in the source code of this
# section for additional configurations
```


#### Map 3: Interactive point map with zoom and hover text  {#map3 .tabset .tabset-pills}

Map 3 shows another example of an interactive map that plots specific points instead of using a shapefile. This map uses a data-set containing information on the location of benefits offices (JBO) and the number of available vacancies.

The location points are displayed on the map and JBO names as well as available vacancies show on hover or click. This interactive map also includes zoom functionality and a view reset button. Points on the map can be resized according to the data or also grouped together. 

The data underlying the map is available to view and download in a tab for users who cannot perceive images. 

##### Map (interactive) output

```{r map_interactive_2, echo = TRUE, warnings = FALSE, message = FALSE, fig.alt="alt text for interactive map", fig.cap="caption for interactive map", out.width="100%"}
library(leaflet)

#### interactive map ####

map3_interactive_map <- leaflet(options = leafletOptions(
  minZoom = 8,
  maxZoom = 12
)) %>%
  setView(lng = -6.5, lat = 54.6, zoom = 8.15, ) %>%
  addProviderTiles("Esri.WorldGrayCanvas")


map3_interactive_map <- addCircleMarkers(
  map = map3_interactive_map, data = df_jbo, lng = ~Longitude, lat = ~Latitude,
  color = "blue",
  radius = 10,
  label = lapply(paste0(
    df_jbo$JBO, ": ", df_jbo$Vacancies,
    " vacancies"
  ), HTML),
  labelOptions = labelOptions(style = list("font-size" = "14px"))
) %>%
  addEasyButton(easyButton(
    icon = "fa-globe",
    title = "Reset Zoom",
    onClick = JS("function(btn, map){ map.setView([54.6,lng = -6.5],8.15); }")
  ))


map3_interactive_map





df_jbo <- df_jbo[1:2]
```


```{r}
f_make_tables(
  data = df_jbo,
  title = "Map3_Data",
  data_style = ns_comma,
  data_dir = paste0(here(), "/code/demo/demo_outputs/", "figdata/")
)
```


##### Map (interactive) data

```{r map_interactive_2b, echo = TRUE}
kable(df_jbo)
```

```{r}
f_make_tables(
  data = df_jbo,
  title = "Map3_Data",
  data_style = ns_comma,
  data_dir = paste0(here(), "/code/demo/demo_outputs/", "figdata/")
)
```


```{r alt text for maps, echo = FALSE}
# Alt text values for tabRemove() function in the interactive map
map_alt <- paste0(
  df_map_data$lgd2014name, " ",
  df_map_data$map_label_value_only
)
map_alt <- paste(map_alt, collapse = '", "')
map_alt <- paste0('["', map_alt, '"]')
```  

#### {-}
